---
title: "p8105_hw1_jf3200"
author: "Jessica Flynn"
date: "9/14/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1 

Here, we construct a data frame including one variable of each type: numeric, logical, character and factor

```{r setup_df}
library(tidyverse)

# Creat the data frame as per instructions

df = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_fact = factor(c("one", "two", "three", "one", "two", "one", "three", "three", "two", "one"))
)
```


We will try to take the means of each of the variables in our data frame.We can see that the means cannot be computed for character or factor variables as we get a warning below

```{r means}

# Try to take the mean of each variable in the data frame

mean(pull(df, norm_samp))
mean(pull(df, norm_samp_pos))
mean(pull(df, vec_char))
mean(pull(df, vec_fact))

```

Next, we will try to coerce the variables to numeric variables using the `as.numeric` function and then try to compute the mean. The output for this is not shown. We observe that the character and factor variables still produce a warning and no mean is calculated (NA returned), however, the logical variable is converted to numeric with a mean of `r mean(pull(df, norm_samp_pos))`. The FALSE becomes 0 and the TRUE becomes 1 in order to compute the mean.  . 

```{r convert_to_numeric, results=FALSE}

#Try to convert each variable to numeric and take the mean

mean(pull(df, as.numeric(norm_samp_pos)))
mean(pull(df, as.numeric(vec_char)))
mean(pull(df, as.numeric(vec_fact)))


```


Below, we will try coering the logical vector to numeric and multiplying by the numeric random sample. Then, we will repeat the steps by coering it to a factor, and also to a factor and then to a numeric in one step. 

We get an error when we try to multiple the logical variable transformed into a factor variable by the numeric variable. We do not get an error for the other conversions. 

```{r convert_and_multiply}
#convert the logical vector to numeric, and multiply the random sample by the result

as.numeric(pull(df,norm_samp_pos)) * pull(df, norm_samp)

#convert the logical vector to a factor, and multiply the random sample by the result

as.factor(pull(df,norm_samp_pos)) * pull(df, norm_samp)

#convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

as.numeric(as.factor(pull(df,norm_samp_pos))) * pull(df, norm_samp)

```

