---
title: "p8105_hw1_jf3200"
author: "Jessica Flynn"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Problem 1 

Here, we construct a data frame including one variable of each type: numeric, logical, character and factor

```{r setup_df}
# Create the data frame as per instructions

prob1_df = 
  tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_fact = factor(c("one", "two", "three", "one", "two", "one", "three", "three", "two", "one"))
  )
```


We will try to take the means of each of the variables in our data frame.We can see that the means cannot be computed for character or factor variables as we get a warning below and an output of NA

```{r means}
# Try to take the mean of each variable in the data frame
mean(pull(prob1_df, norm_samp))
mean(pull(prob1_df, norm_samp_pos))
mean(pull(prob1_df, vec_char))
mean(pull(prob1_df, vec_fact))
```

Next, we will try to coerce the variables to numeric variables using the `as.numeric` function. The output for this is not shown. 

We observe that the logical vector is able to be converted to numeric. The FALSE becomes 0 and the TRUE becomes 1 in order to compute the mean. 

The character vector is converted to all NA, which explains why the mean was NA when we took it previously. 

The factor variable is converted to numbers 1,2 and 3. The numbers are assigned based on alphabetical order of the factor variable. Previously, we were unable to get a mean for the factor variable since it was not numeric or logical, but when we convert it to numeric we will be able to get a mean. 

```{r convert_to_numeric, results=FALSE}
#Try to convert each variable to numeric 
as.numeric(pull(prob1_df, norm_samp_pos))
as.numeric(pull(prob1_df, vec_char))
as.numeric(pull(prob1_df, vec_fact))
```


Below, we will try coercing the logical vector to numeric and multiplying by the numeric random sample. Then, we will repeat the steps by coercing it to a factor, and also to a factor and then to a numeric in one step. 

We get a warning and NAs when we try to multiple the logical variable transformed into a factor variable by the numeric variable. We do not get an error for the other conversions and we get an output of numbers

```{r convert_and_multiply}
#convert the logical vector to numeric, and multiply the random sample by the result

as.numeric(pull(prob1_df,norm_samp_pos)) * pull(prob1_df, norm_samp)

#convert the logical vector to a factor, and multiply the random sample by the result

as.factor(pull(prob1_df,norm_samp_pos)) * pull(prob1_df, norm_samp)

#convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

as.numeric(as.factor(pull(prob1_df,norm_samp_pos))) * pull(prob1_df, norm_samp)

```

## Problem 2 

```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

The penguins dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. The variables contained in the dataset are `r colnames(penguins)`. The variables `sex`, `species` and `island` have class `r class(pull(penguins, sex))`.  The variables  `bill_length_mm` and  `bill_dept_mm` have class `r class(pull(penguins, bill_depth_mm))` while the variables  `flipper_length_mm`, `body_mass_g` and `year` have class `r class(pull(penguins, year))`. 

The mean of flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm=TRUE)` excluding the NA entries. 


We will show a scatterplot of `flipper_length_mm` vs. `bill_length_mm` that is color coded by `species`

```{r scatterplot}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

#export to project directory
ggsave(filename = "hw1_scatterplot_bill_by_flipper_length.png")

```

Looking at the plot, we see that the Gentoo species has the largest flipper length on average. The flipper length of Adelie and Chinstrap are similar in size. We also see that the Chinstrap and Gentoo species have the largest bill lengths, followed by the Adelie. There appears to be a slight positive correlation between bill length and flipper length among all species. 